---
title: "HW01"
output: html_document
date: "2024-04-25"
---


```{r setup}
### Edit this line so your notebook renders properly
knitr::opts_knit$set(root.dir = normalizePath("~/KGI Spring 2024/GENE 5130 - Bioinformatics in R/R files")) 

```

```{r}
### Import the important libraries
library(WGCNA)
library(dplyr)
library(readr)
library(DESeq2)
library(ggplot2)
library(org.Hs.eg.db)
library(gprofiler2)
library(clusterProfiler)

```

```{r}
### Run this chunk to import the counts table and metadata into your environment
counts <- read.csv('../data/GSE152418RawCounts.csv', header = TRUE, row.names = 1)
metadata <- read.csv('../data/GSE152418Metadata.csv', header = TRUE, row.names = 1)

```

```{r}
gene2color <- read.csv('../data/Gene2color-1.csv', header = TRUE)
p_values <- read.csv('../data/Pvalues-1.csv', header = TRUE)
correlations <- read.csv('../data/Correlations-1.csv', header = TRUE)
  
```

```{r}
### Subset the data so only the genes that passed the filter are kept
subset <- goodSamplesGenes(t(counts))
good_gene_counts <- counts[subset$goodGenes,]

### Filter the data to remove bad samples
better_gene_counts <- good_gene_counts %>%
  dplyr::select(-GSM4614993) %>% 
  dplyr::select(-GSM4614995) %>% 
  dplyr::select(-GSM4615000)

```
Run Gene Set Enrichment Analysis / Gene Ontology on a set of genes (color) <- turquoise

```{r}
turquoise_genes <- gene2color[gene2color$COLOR == "turquoise",]

```

```{r}
turq_gene_counts <- better_gene_counts[(row.names(better_gene_counts) %in% turquoise_genes$GENEID),]

```

```{r}
gsm_list <- c('GSM4614993', 'GSM4614995', 'GSM4615000')
new_metadata <- metadata[!(row.names(metadata) %in% gsm_list),]

temp_rename <- rename
rename <- dplyr::rename

new_metadata %>% 
  rename(days_post_symptom_onset = days_post_symptom_onset.ch1) %>%
  rename(disease_state = disease.state.ch1) %>%
  rename(gender = gender.ch1) %>%
  rename(geographical_location = geographical.location.ch1) %>%
  rename(severity = severity.ch1)

```

```{r}
### Create a deseq object
dds <- DESeqDataSetFromMatrix(countData = turq_gene_counts, 
                              colData = new_metadata,
                              design = ~ 1)
```

```{r}
dds75 <- dds[rowSums(counts(dds)) >= 23]

```

```{r}
### Run deseq2
deseq_ob <- DESeq(dds75)

### Save the results to a new object
res <- results(deseq_ob, alpha = 0.05)

```

```{r}
### Convert between ENSEMBL ID and symbol (gene names) 
sigs.df <-  as.data.frame(res)
sigs.df$symbol <- mapIds(org.Hs.eg.db, keys= rownames(sigs.df), keytype = 'ENSEMBL', colum = "SYMBOL")

```

```{r}
### Gene Ontology Analysis
diff.df <- as.data.frame(sigs.df)
diff.df <- diff.df %>%
  filter(padj < 0.05, log2FoldChange > 2, baseMean > 100)

genes_to_test <- rownames(diff.df)
GO_results <- enrichGO(gene = genes_to_test, OrgDb = "org.Hs.eg.db", keyType = 'ENSEMBL', ont = "BP")
fit <- plot(barplot(GO_results, showCategory=10))

```

```{r}
go_upreg <- gost(genes_to_test, organism = 'hsapiens', significant = T)
gostplot(go_upreg, interactive = T)

```

```{r}
### GSEA Plot
res <- res[order(-res$stat),]
gene_list <- res$stat
names(gene_list) <- rownames(res)
gse <- gseGO(gene_list, ont = "BP", keyType = "ENSEMBL", OrgDb ="org.Hs.eg.db", eps = 1e-300 )
gseaplot(gse, geneSetID = "GO:0006119" )

```
